The contact value approximation for the pair distribution function of the inhomogeneous hard-sphere fluid

### 2014-05-20 papers/pair-correlation/paper.tex
try to clarify wording of section IV and paragraph vefore Tested on bennet. Took 0:17.17.

Sokolowski and Fischer addressed the shortcomings of the theory of
Fischer and Methfessel by modifying this approach to use
averages, centered on the two points $\rr_1$ and $\rr_2$:
\begin{align}
  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12},
  \tfrac12(\bar{n}(\rr_1)+\bar{n}(\rr_2))\right)
  \label{eq:sokolowski}
\end{align}
where their averaged density $\bar{n}(\rr)$ is given by:
\begin{align}
  \bar{n}(\rr) \equiv \frac{3}{4\pi (0.8\sigma)^3}\int n(\rr')\Theta(0.8\sigma - |\rr-\rr'|) d\rr'
\end{align}
is the density averaged over a sphere with diameter
$0.8\sigma$~\cite{sokolowski1992role, wadewitz2000application, winkelmann2001liquid}.  The
value 0.8 in this formula was arrived at by fitting to Monte Carlo
simulation.  Although Eq.~\ref{eq:sokolowski} has the advantages of
only involving density averages at the points at which the pair
distribution function is desired, it cannot be written as a
single-site convolution, since the convolution kernel depends on both
points.

In a previous paper~\cite{schulte2012using}, we introduced a
functional that gives a good approximation for the pair distribution
function averaged over positions $\rr_2$ that are in contact with
$\rr_1$, defined as:
\begin{align}
  g_\sigma(\rr_1) &\equiv \frac{ \int g^{(2)}(\rr_1,\rr_2) \delta(\sigma -|\rr_1-\rr_2|)n(\rr_2)
    d\rr_2 }{ \tilde{n}(\rr_1)  }
\end{align}
where the weighted density $\tilde{n}(\rr_1)$ is defined by:
\begin{align}
  \tilde{n}(\rr) &\equiv \int n(\rr') \delta(\sigma -|\rr - \rr'|)d\rr'.
\end{align}
We use the contact-value theorem to derive the exact formula:
\begin{align}
  g_\sigma(\rr)% &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
  &= \frac12 \frac{1}{k_BT n(\rr) \tilde{n}(\rr)} \frac{\delta
    F_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:gsigma-exact}
\end{align}
where $\sigma(\rr)$ is the diameter of hard spheres located at
position $\rr$, and $F_{HS}$ is the Helmholtz free energy of the
hard-sphere fluid.  The approximation for $g_\sigma(\rr)$ is obtained
by using the White Bear functional~\cite{roth2002whitebear} for the
Helmholtz free energy $F_{HS}$ in Eq.~\ref{eq:gsigma-exact}.  This
provides an excellent approximation for this averaged value of the
pair distribution function at contact for a variety of interfaces, and
over a wide range of densities.

\section{Contact value approximation}
In the approaches for the pair distribution function mentioned above,
the radial distribution function used in the approximation
was dependent upon the density averaged over some volume.  We seek to
achieve greater accuracy by using instead a function dependent upon
our averaged $g_{\sigma}(\rr)$ discussed above, which holds more
information about an inhomogeneous system than does a simple
convolution of the density.
%
We construct the CVA with the average of two radial distribution
functions, evaluated at the distance between the two points, that are
themselves functions of the averaged pair distribution function at contact
$g_{\sigma}(\rr)$ evaluated at the two points:
%
%% We therefore seek a function form
%% for our radial distribution that is dependent upon $g_{\sigma}(\rr)$
%% and that will approach the correct solution in the homogeneous limit
%% and also in the limit of contact distance between the two points.
\begin{align}
  g^{(2)}(\rr_1,\rr_2) = \frac{g(r_{12}, g_\sigma(\rr_1)) +
    g(r_{12}, g_\sigma(\rr_2))}{2}. \label{eq:g2-our-mean}
\end{align}
This contact value approximation for $g^{(2)}(\rr_1,\rr_2)$ is
constructed to reproduce the exact value for the integral:
\begin{align}
  F_1^{\text{contact}} &= \tfrac12 \iint
  g^{(2)}_{HS}(\rr_1,\rr_2)n(\rr_1)n(\rr_2)\delta(|\rr_1-\rr_2|-\sigma)
  d\rr_1d\rr_2
  \label{eq:mean-field-contact}
\end{align}
which is the mean-field correction to the free energy (see
Eq.~\ref{eq:mean-field}) for a purely contact interaction.

>>>>

Sokolowski and Fischer addressed the shortcomings of the theory of
Fischer and Methfessel by modifying this approach to use
averages, centered on the two points $\rr_1$ and $\rr_2$:
\begin{align}
  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12},
  \tfrac12(\bar{n}(\rr_1)+\bar{n}(\rr_2))\right)
  \label{eq:sokolowski}
\end{align}
where their averaged density $\bar{n}(\rr)$ is given by:
\begin{align}
  \bar{n}(\rr) \equiv \frac{3}{4\pi (0.8\sigma)^3}\int n(\rr')\Theta(0.8\sigma - |\rr-\rr'|) d\rr'
\end{align}
is the density averaged over a sphere with diameter
$0.8\sigma$~\cite{sokolowski1992role}.  The value 0.8 in this formula
was arrived at by fitting to Monte Carlo simulation.  Although
Eq.~\ref{eq:sokolowski} has the advantages of only involving density
averages at the points at which the pair distribution function is
desired, it cannot be written as a single-site convolution, since the
convolution kernel depends on both points.  However, this approach has
been used by two papers to do
\fixme{what?}~\cite{wadewitz2000application, winkelmann2001liquid}.

In a previous paper~\cite{schulte2012using}, we introduced a
functional that gives a good approximation for the pair distribution
function averaged over positions $\rr_2$ that are in contact with
$\rr_1$, defined as:
\begin{align}
  g_\sigma(\rr_1) &\equiv \frac{ \int g^{(2)}(\rr_1,\rr_2) \delta(\sigma -|\rr_1-\rr_2|)n(\rr_2)
    d\rr_2 }{ \tilde{n}(\rr_1)  }
\end{align}
where the weighted density $\tilde{n}(\rr_1)$ is defined by:
\begin{align}
  \tilde{n}(\rr) &\equiv \int n(\rr') \delta(\sigma -|\rr - \rr'|)d\rr'.
\end{align}
We use the contact-value theorem to derive the exact formula:
\begin{align}
  g_\sigma(\rr)% &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
  &= \frac12 \frac{1}{k_BT n(\rr) \tilde{n}(\rr)} \frac{\delta
    F_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:gsigma-exact}
\end{align}
where $\sigma(\rr)$ is the diameter of hard spheres located at
position $\rr$, and $F_{HS}$ is the Helmholtz free energy of the
hard-sphere fluid.  The functional derivative with of free energy with
respect to hard-sphere diameter in Eq.~\ref{eq:gsigma-exact} requires
that we be able to evaluate the change in free energy resulting from a
change in the diameter of specifically the hard spheres located at
position $\rr$.  This somewhat unusual construction is mathematically
straightforward within Fundamental Measure Theory
(FMT)~\cite{rosenfeld1989free}.  We employ the White Bear variation of
the FMT free energy functional~\cite{roth2002whitebear}, which
provides an excellent approximation for this averaged value of the
pair distribution function at contact for a variety of interfaces, and
over a wide range of densities.

\section{Contact value approximation}
In the approaches for the pair distribution function mentioned above,
the radial distribution function used in the approximation
was dependent upon the density averaged over some volume.  We seek to
achieve greater accuracy by making use of a function dependent upon
our averaged $g_{\sigma}(\rr)$ discussed above, which holds more
information about an inhomogeneous system than does a simple
convolution of the density.
%
We construct the CVA with the average of two radial distribution
functions, evaluated at the distance between the two points, that are
themselves functions of the averaged pair distribution function at contact
$g_{\sigma}(\rr)$ evaluated at the two points:
%
%% We therefore seek a function form
%% for our radial distribution that is dependent upon $g_{\sigma}(\rr)$
%% and that will approach the correct solution in the homogeneous limit
%% and also in the limit of contact distance between the two points.
\begin{align}
  g^{(2)}(\rr_1,\rr_2) = \frac{g(r_{12}, g_\sigma(\rr_1)) +
    g(r_{12}, g_\sigma(\rr_2))}{2}. \label{eq:g2-our-mean}
\end{align}
This contact value approximation for $g^{(2)}(\rr_1,\rr_2)$ is
constructed to reproduce the exact value for the integral:
\begin{align}
  F_1^{\text{contact}} &= \tfrac12 \iint
  g^{(2)}_{HS}(\rr_1,\rr_2)n(\rr_1)n(\rr_2)\delta(|\rr_1-\rr_2|-\sigma)
  d\rr_1d\rr_2
  \label{eq:mean-field-contact}
\end{align}
which is the mean-field correction to the free energy (see
Eq.~\ref{eq:mean-field}) for a purely contact interaction.
